---
title: "IMU Objects in the GMWM R Package"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IMU Objects in the GMWM R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction
Within this guide, we consider a real signal issued from a MEMS IMU. 

# Load the `gmwm` package
First, make sure to load the `gmwm` package into R.

```{r load_gmwm}
library("gmwm")
```

# Installation
To obtain the data, we must download and install the `imudata` package.

```{r install_imu, message=FALSE}
if(!require("imudata")){
  install_imudata()
  library("imudata")
}
```

# Loading Data
To load the data and display the first rows for each sensor we do:
```{r imuData, message=FALSE}
data(imu6)
head(imu6)
```

To use the data within the modeling paradigm, we must first cast it as type imu:
```{r imu_object}
sensor = imu(imu6, 
             gyroscope = 1:3,
             accelerometer = 1:3, 
             axis = c('x','y','z'))
```


# Wavelet Variance
To plot the WV (and confidence intervals) for all gyroscopes and accelerometers, we can use the following function 

```{r imuPlot separate}
wv.imu = wvar(sensor)
plot(wv.imu)
```

# Model Estimation
Let us consider the Gyro on the axis X and, given the estimated WV from the plot, let us estimate a model which is the sum of three AR(1) processes:

```{r imuModel}
# Define model
TS.mod.imu = 3*AR1()

# Compute GMWM estimator
model.imu = gmwm.imu(TS.mod.imu, data = imu6[,1])
```
**Note: The gmwm.imu function only is able to handle one sensor at a time.**

To see the results of the estimation, we can use the summary function as well as visualize the fit of the implied WV to the estimated WV from Gyro X (as done in the simulation example).

```{r imuModel summary}
summary(model.imu)
plot(model.imu)
```